---
title: "HW4"
author: "YIXI XU"
date: "11/4/2023"
output: pdf_document
---

**1.** Create a vector containing each number between 1 and 50, inclusively.
```{r}
v <- c(1:50)
```

a. Use your vector to construct a 5 x 10 matrix.
```{r}
m <- matrix(v,5,10,byrow=T)
m
```

b. Create a logical matrix where the values of the matrix is evenly divisible by 3 or has a remainder of 1 when divided by 6.
```{r}
lm <- (m %% 3 == 0) | (m %% 6 == 1)
lm
```

c. Extract the values of the matrix where the logical matrix made in part b.
```{r}
extract <- m[lm]
extract
```

d. Create a function called hailstone(x). The function should determine if x is odd or even. If x is odd, multiply x by 3 and add 1, otherwise divide x by 2. Repeat this process until the number is 1. The function should return the number of the cycles that the number goes through before reaching 1.
Apply the function that you created in part d to the values extracted in part c.
```{r}
hailstone <- function(x) {
  s <- 0 
  while (x != 1) {
     if (x %% 2 == 1) {
      x <- 3 * x + 1
    } else {
      x <- x / 2
    }
    s <- s + 1
  }
  return(s)
}
sapply(extract, hailstone)
```


**2.** Install the ISLR library and load the Auto dataset.
```{r}
if (!require(ISLR, quietly = TRUE)) {
  install.packages("ISLR")
}
library(ISLR)
data(Auto)
str(Auto)
```

a.Remove qualitative data columns from the data.
```{r}
auto1 <- Auto[,sapply(Auto,is.numeric)]
str(auto1)
```

b.Write a loop that will iterate through the columns and print the range of each quantitative variable.
```{r}
for (j in names(auto1)) {
  min <- min(auto1[,j])
  max <- max(auto1[,j])
  cat(j,"is from", min, "to", max, "\n")
}
```

c.Create a matrix containing the mean and the standard deviation of each column.
```{r}
mat1<-matrix(c(colMeans(auto1), sapply(auto1, sd)),ncol=2)
colnames(mat1) <- c("Mean", "SD")
rownames(mat1) <- c(names(auto1))
mat1
```

d.Identify rows that have less than the mean amount of mpg. Remove those rows from the dataset.
```{r}
auto2 <- auto1[auto1$mpg >= mean(auto1$mpg), ]
dim(auto2)
```

e.Recalculate the mean and standard deviation of each column now that values with below average mpg have been removed.
```{r}
mat2<-matrix(c(colMeans(auto2), sapply(auto2, sd)),ncol=2)
colnames(mat2) <- c("Mean", "SD")
rownames(mat2) <- c(names(auto2))
mat2
```

**3.**  Comparison

It's easier to create matrix and peform mean/var calculation with R than Python. In python you need to use libraries such as NumPy/Pandas
However, it's easier and straightforward to do data manipulation with python such as compute mean/sd in each column of the dataset (with simple syntax command).
For basic and tradtional statistical task i would perfer R but in some machine learning task such as neuro network iw would use python.

